---
title: "PC Unlimited Answers"
output:
  pdf_document: default
  html_notebook: default
  html_document:
    df_print: paged
---

```{r, include=FALSE}
library(tidyverse) #a handy tool for basic data analysis
library(gmodels) # library for the command "CrossTable"


# set working directory, for example, it can be where you save the data, 
# or where you plan to save the output
# setwd("~/Downloads")
url <- "https://tinyurl.com/pcunlimited-csv"
PCUnlimited <- read.csv(url) #load the data
```


```{r}
# Take a peek at the data
head(PCUnlimited)
tail(PCUnlimited)
```

****** Q1, What percent of PCUnlimited customers are female?
```{r}
CrossTable(PCUnlimited$gender)

# An alternative way, dplyr, which is much more versatile
PCUnlimited %>% 
  group_by(gender) %>% 
  summarise(avg_spend_by_gender=mean(totdol), num_of_customer_by_gender=n()) %>% 
  mutate(percentage=num_of_customer_by_gender/sum(num_of_customer_by_gender))
```
****** Q2, which regions has the highest number of customers?
```{r}
table(PCUnlimited$region)
sort(table(PCUnlimited$region), decreasing = TRUE)
# using pipe from dplyr
PCUnlimited %>% 
  group_by(region) %>% 
  summarise(num_of_customer_by_region=n()) %>% 
  arrange(desc(num_of_customer_by_region))
```

****** Q3, which regions have the highest and lowest average total spendings?
```{r}
PCUnlimited %>% 
  group_by(region) %>% 
  summarise(avg_spend_by_region=mean(totdol)) %>% 
  arrange(desc(avg_spend_by_region))
```

****** Q4, What are the average $ spent and the total # of purchases by region? 
```{r}
PCUnlimited %>% group_by(region) %>% 
  summarise(avg_spending = mean(totdol), tot_orders = sum(numpurch))
```

****** Q5, What are the average $ spent and the total # of purchases by gender? 
```{r}
PCUnlimited %>% 
  group_by(gender) %>% 
  summarise(avg_spend_by_gender = mean(totdol), 
            tot_orders_by_gender = sum(numpurch))
```

****** Q6, Calculate the correlation between total spending and total # of orders. 
```{r}
cor(PCUnlimited$totdol, PCUnlimited$numpurch)
cor.test(PCUnlimited$totdol, PCUnlimited$numpurch)
```

****** Q7, what are the percentages of male and female who responded to the offer?
```{r}
response_by_gender <- PCUnlimited %>% 
                      group_by(gender, respond) %>% 
                      summarise(n_by_gender_response=n()) %>% 
  mutate(percentage = n_by_gender_response/sum(n_by_gender_response)*100)
```

****** Q8, what are min, max, and average of individual spending per order?
```{r}
spending_per_order <- PCUnlimited %>% 
                        mutate(spending_per_order=totdol/numpurch) %>% 
                        summarise(min=min(spending_per_order), 
                                  max=max(spending_per_order),
                                  avg=mean(spending_per_order))
```

****** Q9, can we show a bar chart of q3 above?
```{r}
spending_by_region <- PCUnlimited %>% 
                        group_by(region) %>% 
                        summarise(average_spending=mean(totdol)) %>% 
                        arrange(desc(average_spending))
bar_spending <- 
  ggplot(data=spending_by_region, 
         aes(x = region, y = average_spending)) + 
  labs(x="Region", y="Average Spending") + 
  geom_bar(stat="identity", fill="steelblue")
bar_spending

# without stat="identity", bar chart shows "count", 
# which is one for each region in this case
ggplot(data=spending_by_region, 
       aes(x = region)) + 
  labs(x="Region", y="Average Spending") + 
  geom_bar(fill="steelblue") 

# how about sort the region by their average spending?
bar_spending_sorted <- 
  ggplot(data=spending_by_region, 
         aes(x = reorder(region, -average_spending), y = average_spending)) + 
  labs(x="Region", y="Average Spending") + 
  geom_bar(stat="identity")
bar_spending_sorted
```

****** Q10. Create a new dataframe that contains only the variables of cust_num, respond, 
and gender
```{r}
PCUlimited_subset <- PCUnlimited %>% 
  select(cust_num, respond, gender)
head(PCUlimited_subset)
```

****** Q11. Create a new dataframe that contains only urban female customers
```{r}
PCUnlimited_urban_female <- PCUnlimited %>% 
  filter(gender=="female" & addresstype=="urban")
head(PCUnlimited_urban_female)
```

